\section{Definition and basic properties of a CW-complex}

We choose to formalise the historical definition because the modern one would require us to consider the $n$-skeletons as topological spaces which would create a lot of work unifying them to work well together. 
The historical definition, however, allows us to work in one topological space which avoids this issue.  

The following is our definition of CW-complexes in Lean: 
\href{https://github.com/scholzhannah/CWComplexes/blob/7be4872a05b534011cc969eb5b80a4b7f0bf57e2/CWcomplexes/Definition.lean#L47-L76}{\faExternalLink}

\begin{lstlisting}
class CWComplex.{u} {X : Type u} [TopologicalSpace X] (C : Set X) where
  cell (n : ℕ) : Type u
  map (n : ℕ) (i : cell n) : PartialEquiv (Fin n → ℝ) X
  source_eq (n : ℕ) (i : cell n) : (map n i).source = closedBall 0 1
  cont (n : ℕ) (i : cell n) : ContinuousOn (map n i) (closedBall 0 1)
  cont_symm (n : ℕ) (i : cell n) : ContinuousOn (map n i).symm (map n i).target
  pairwiseDisjoint' :
    (univ : Set (Σ n, cell n)).PairwiseDisjoint (fun ni ↦ map ni.1 ni.2 '' ball 0 1)
  mapsto (n : ℕ) (i : cell n) : ∃ I : Π m, Finset (cell m),
    MapsTo (map n i) (sphere 0 1) (⋃ (m < n) (j ∈ I m), map m j '' closedBall 0 1)
  closed' (A : Set X) (asubc : A ⊆ C) : IsClosed A ↔ ∀ n j, IsClosed (A ∩ map n j '' closedBall 0 1)
  union' : ⋃ (n : ℕ) (j : cell n), map n j '' closedBall 0 1 = C
\end{lstlisting}

The \lstinline|.{u}| is a way to fix a universe level so that our definition of a CW-complex does not depend on a number of different universe levels: The one of \lstinline{X} and the one of \lstinline{cell n} for every $n \in \bN$.
\lstinline{cell (n : ℕ)} represents the indexing set that we called $I_n$ in Definition \ref{defi:CWComplex2}. \lstinline{map (n : ℕ) (i : cell n)} represent what we called $Q_i^n$ in that definition.

\lstinline{Fin n}
\href{https://github.com/leanprover/lean4/blob/0edf1bac392f7e2fe0266b28b51c498306363a84/src/Init/Prelude.lean#L1826-L1838}{\faExternalLink} 
is the set containing $n$ natural numbers starting at 0.
\lstinline{Fin n → ℝ} is one way to express $\bR^n$ in Lean.
\lstinline{PartialEquiv} is a structure defined in mathlib as follows: 
\href{https://github.com/leanprover-community/mathlib4/blob/ed125a4216d18273cb1b96d4c846d32b85d74faf/Mathlib/Logic/Equiv/PartialEquiv.lean#L113-L134}{\faExternalLink}

\begin{lstlisting}
structure PartialEquiv (α : Type*) (β : Type*) where
  toFun : α → β
  invFun : β → α
  source : Set α
  target : Set β
  map_source' : ∀ ⦃x⦄, x ∈ source → toFun x ∈ target
  map_target' : ∀ ⦃x⦄, x ∈ target → invFun x ∈ source
  left_inv' : ∀ ⦃x⦄, x ∈ source → invFun (toFun x) = x
  right_inv' : ∀ ⦃x⦄, x ∈ target → toFun (invFun x) = x
\end{lstlisting}

It bundles two maps and two sets that get mapped to each other by the respective maps.
Restricting the maps to these sets yields two maps that are the inverse of each other.
We use this instead of a similar construction called \lstinline{Equiv} for bijections to avoid explicitly having to deal with restrictions.
The brackets \lstinline{⦃⦄} are similar to the curly brackets and are used here since \lstinline{x} can be inferred from the left sides of the implications.

The property \lstinline{source_eq} specifies the source of the \lstinline{PartialEquiv}.
\lstinline{cont} and \lstinline{cont_symm} make the bijection into a homeomorphism giving us property (i) of Definition \ref{defi:CWComplex2}.
The property \lstinline{pairwiseDisjoint'} corresponds to property (ii) of Definition \ref{defi:CWComplex2}.
We are adding the prime to its name because we will later see a lemma called \lstinline{pairwiseDisjoint} that we prefer to use.
\lstinline{fun a : α ↦ f a} for \lstinline{f : α → β} is a way to construct a map.

\lstinline{mapsto} is the equivalent of property (iii) of the definition of a CW-complex.
The \lstinline{Π} defines a dependent function type which we discussed in Section \ref{sec:typetheory}.
\lstinline{Finset α}
\href{https://github.com/leanprover-community/mathlib4/blob/ed125a4216d18273cb1b96d4c846d32b85d74faf/Mathlib/Data/Finset/Basic.lean#L133-L139}{\faExternalLink}
is the type of all finite sets in a type \lstinline{α}.
It can be imagined as a set bundled with the information that it is finite (but note that the actual definitions of \lstinline{Finset α} and \lstinline{Set α} are quite different).
\lstinline{MapsTo} is defined in mathlib as
\href{https://github.com/leanprover-community/mathlib4/blob/ed125a4216d18273cb1b96d4c846d32b85d74faf/Mathlib/Data/Set/Operations.lean#L231-L232}{\faExternalLink}

\begin{lstlisting}
def MapsTo (f : α → β) (s : Set α) (t : Set β) : Prop :=
  ∀ ⦃x⦄, x ∈ s → f x ∈ t
\end{lstlisting}

and is relatively self-explanatory.  

\lstinline{closed'} represents property (iv) of Definition \ref{defi:CWComplex2} and \lstinline{union'} represents property (v).

\medskip

We have chosen this to be a class so that we can make use of typeclass inference which we explained in Section \ref{sec:implicitandtypeclass}.

There are a few things to note about this formalisation of the definition.
First of all it does not require $X$ to be a Hausdorff space.
This is done so that when you define a CW-complex you can choose to first define the structure in this way and later show that it is a Haussdorff space to apply lemmas about CW-complexes of which almost all will require that $X$ is Hausdorff.
Additionally we introduce a relative component:
Instead of defining what it means for a space to be a CW-complex we define what it means for a subspace $C$ of $X$ to be a CW-complex in $X$.
This is useful firstly to be able to work with a nicer topology:
If you consider $S^1$ as a CW-complex and a subspace of $\bR$ you might find it easier to work with the topology on $\bR$ instead of the subspace topology.
Secondly constructions such as attaching cells or taking disjoint unions of CW-complexes might be easier to work with if you are already working in the same overarching type.
This approach is inspired by \cite{Gonthier2013} where the authors notice that it is helpful to consider subsets of an ambient group to avoid having to work with different group operations and similar issues.

One question that naturally arises is whether these changes to the definition preserve the notion of a CW-complex.
Firstly note that if we choose $X$ and $C$ to be the same we recover Definition \ref{defi:CWComplex2} exactly.
Now let us think about what happens if we choose $X$ and $C$ to be different.
Firstly this allows us to conclude that $C$ is closed:

\begin{lem} \label{lem:Cclosed}
  Let $X$ be a Hausdorff space and and $C$ a CW-complex in $X$ as in the formalised definition.
  Then $C$ is closed.
  \href{https://github.com/scholzhannah/CWComplexes/blob/7be4872a05b534011cc969eb5b80a4b7f0bf57e2/CWcomplexes/Definition.lean#L234-L238}{\faExternalLink}
\end{lem}
\begin{proof}
  Since $C \subseteq C$ it is enough to show that $C \cap \closedCell{n}{i}$ is closed for every $n \in \bN$ and $i \in I_n$.
  But by the property \lstinline{union'} we know that $C \cap \closedCell{n}{i} = \closedCell{n}{i}$ which is closed by the same argument as in the proof of Lemma \ref{lem:closedCellclosed}.
\end{proof}

This indeed excludes some CW-complexes:

\begin{example}
  Let $I \subseteq \bR$ be an open interval.
  Then $I$ is a CW-complex in the sense of Definition \ref{defi:CWComplex2}.
\end{example}
\begin{proof}
  Since $I$ is homeomorphic to $\bR$ it is by Lemma \ref{lem:homeomorph} enough to show that $\bR$ admits the structure of a CW-complex.
  As $0$-cells we choose every $z \in \bZ \subseteq \bR$.
  As $1$-cells we choose the intervals $(z, z + 1)$ for every $z \in \bZ \subseteq \bR$.
  Properties (i), (ii), (iii), and (v) of Definition \ref{defi:CWComplex2} are easy to verify.
  We will therefore focus on property (iv), i.e. the weak topology.
  The forward implication follows in the same manner as in a lot of other proofs.
  Let us thus move on to the backwards direction.
  Take $A \subseteq \bR$ and assume that $A \cap [z, z + 1]$ is closed for all $z \in \bZ$.
  We now need to show that $A$ is closed.
  It is wellknown that $\bR$ is a metric space and by Lemma \ref{lem:metricissequential} it is in particular a sequential space.
  It is therefore enough to show that for every convergent sequence $(a_n)_{n \in \bN} \subseteq A$ the limit point is also in $A$.
  Take an arbitrary convergent $(a_n)_{n \in \bN} \subseteq A$.
  We call the limit point $a$.
  Then there exists a $z \in \bZ$ such that $a \in (z, z + 2)$.
  Thus there is a subsequence $(a'_n)_{n \in \bN} \subseteq A \cap [z, z + 2]$ which obviously also converges to $a$.
  But by assumption $A \cap [z, z + 2] = (A \cap [z, z + 1]) = (A \cap [z + 1, z + 2])$ is closed and therefore sequentially closed which gives us that $a \in A \cap [z, z + 2] \subseteq A$.
\end{proof}

But remember that our definition in Lean still allows us to view an open interval as a CW-complex in itself.

And every space that fulfils the formalised definition also fulfils Definition \ref{defi:CWComplex2}:

\begin{lem}
  Let $C$ be a CW-complex in a Hausdorff space $X$ as in the definition in the formalisation.
  Then $C$ is a CW-complex as in Definition \ref{defi:CWComplex2}.
\end{lem}
\begin{proof}
  Properties (i), (ii), (iii) and (v) of Definition \ref{defi:CWComplex2} are immediate.
  Thus let us look at property (iv).
  We assume that
  \[A \subseteq C \text{ is closed in } X \iff \closedCell{n}{i} \cap A \text{ is closed in } X \text{ for all } n \in \bN \text{ and } i \in I_n\]
  and need to show that
  \[A \subseteq C \text{ is closed in } C \iff \closedCell{n}{i} \cap A \text{ is closed in } C \text{ for all } n \in \bN \text{ and } i \in I_n.\]
  It is easy to see that the forward direction is true.
  For the backwards direction take $A \subseteq C$ such that $A \cap \closedCell{n}{i}$ is closed in $C$ for all $n \in \bN$ and $i \in I_n$.
  That means that for every $n \in \bN$ and $i \in I_n$ there is a closed set $B_i^n \subseteq X$ such that $B^n_i \cap C = A \cap \closedCell{n}{i}$.
  But since $C$ is closed by Lemma \ref{lem:Cclosed} that means that $A \cap \closedCell{n}{i}$ was already closed for every $n \in \bN$ and $i \in I_n$.
  Thus we are done by assumption.
\end{proof}

With that we can move onto the last important difference that our new definition has.
While \lstinline{Fin n → ℝ} is a way to represent $\bR^n$ in Lean it does not actually carry the euclidean metric but the maximum metric.
So instead of considering closed balls we are looking at cubes which does not change our definition since the two are homeomorphic.
We could use the euclidean metric on $\bR^n$ which would be written as \lstinline{EuclideanSpace ℝ (Fin n)} but since we are mostly arguing abstractly about CW-complexes this is unnecessary and takes up more space.

When proving that something is a CW-complex in Lean it can often be surprising how much longer and more technical the formalised proofs of closure finiteness, i.e. the property \lstinline{mapsto}, can be. 
One reason for this is that our proofs in the first chapter heavily rely on our intuition about finiteness while Lean obligates us to show finiteness of the union explicitly and in detail.

The code in the rest of the section will always have the following assumptions:

\begin{lstlisting}
variable {X : Type*} [t : TopologicalSpace X] [T2Space X] {C : Set X} [CWComplex C]
\end{lstlisting}

where \lstinline{T2Space X}
\href{https://github.com/leanprover-community/mathlib4/blob/ed125a4216d18273cb1b96d4c846d32b85d74faf/Mathlib/Topology/Separation.lean#L1291-L1297}{\faExternalLink}
expresses that \lstinline{X} is a Hausdorff space.

We also want to define some notation in Lean.
Just as we defined $\closedCell{n}{i}$ to represent $Q^n_i(D_i^n)$ and similar notations in Definition \ref{defi:CWComplex2} we can do the same in the formalisation: 
\href{https://github.com/scholzhannah/CWComplexes/blob/7be4872a05b534011cc969eb5b80a4b7f0bf57e2/CWcomplexes/Definition.lean#L84-L94}{\faExternalLink}

\begin{lstlisting}
def openCell (n : ℕ) (i : cell C n) : Set X := map n i '' ball 0 1

def closedCell (n : ℕ) (i : cell C n) : Set X := map n i '' closedBall 0 1

def cellFrontier (n : ℕ) (i : cell C n) : Set X := map n i '' sphere 0 1
\end{lstlisting}

We can now state some of the properties of our definition with this new notation.
We restate \lstinline{pairwiseDisjoint'} in two ways:
\href{https://github.com/scholzhannah/CWComplexes/blob/7be4872a05b534011cc969eb5b80a4b7f0bf57e2/CWcomplexes/Definition.lean#L96-L103}{\faExternalLink}

\begin{lstlisting}
lemma pairwiseDisjoint :
    (univ : Set (Σ n, cell C n)).PairwiseDisjoint (fun ni ↦ openCell ni.1 ni.2) := ⋯

lemma disjoint_openCell_of_ne {n m : ℕ} {i : cell C n} {j : cell C m}
    (ne : (⟨n, i⟩ : Σ n, cell C n) ≠ ⟨m, j⟩) :
    openCell n i ∩ openCell m j = ∅ := ⋯
\end{lstlisting}

The second one is especially convenient to use as the hypothesis \lstinline{ne} can often be automatically verified by a tactic called \lstinline{aesop}.
Information on \lstinline{aesop} can be found in \cite{Limperg2023}.

The properties \lstinline{closed'} and \lstinline{union'} can be rewritten with the new notation as follows:
\href{https://github.com/scholzhannah/CWComplexes/blob/7be4872a05b534011cc969eb5b80a4b7f0bf57e2/CWcomplexes/Definition.lean#L112-L116}{\faExternalLink}

\begin{lstlisting}
lemma closed (A : Set X) (asubc : A ⊆ C) :
  IsClosed A ↔ ∀ n (j : cell C n), IsClosed (A ∩ closedCell n j) := ⋯

lemma union : ⋃ (n : ℕ) (j : cell C n), closedCell n j = C := ⋯
\end{lstlisting}

As in Definition \ref{defi:CWComplex2} we also want to define notation for the $n$-skeletons.
In the first chapter we often chose to start inductions at $-1$ to make the base case trivial.
When formalising we want to be able to use the already defined induction principles that naturally start at $0$.
For that purpose we use an auxiliary definition called \lstinline{levelaux} that is shifted by $1$ in comparison to the usual notion of the $n$-skeleton which we call \lstinline{level}: 
\href{https://github.com/scholzhannah/CWComplexes/blob/7be4872a05b534011cc969eb5b80a4b7f0bf57e2/CWcomplexes/Definition.lean#L137-L146}{\faExternalLink}

\begin{lstlisting}
def levelaux (C : Set X) [CWComplex C] (n : ℕ∞) : Set X :=
  ⋃ (m : ℕ) (_ : m < n) (j : cell C m), closedCell m j

def level (C : Set X) [CWComplex C] (n : ℕ∞) : Set X :=
  levelaux C (n + 1)
\end{lstlisting}

Note that we are choosing \lstinline{n} in \lstinline{ℕ∞} \href{https://github.com/leanprover-community/mathlib4/blob/ed125a4216d18273cb1b96d4c846d32b85d74faf/Mathlib/Data/ENat/Basic.lean#L27-L35}{\faExternalLink} which is the type of natural numbers extended by infinity which can be written as \lstinline{⊤}.
Since \lstinline{level} is defined in terms of \lstinline{levelaux} it is often trivial to derive a lemma about \lstinline{level} from the corresponding lemma about \lstinline{levelaux}.

We can also define what it means for a CW-complex to be finite dimensional, of finite type, or finite:
\href{https://github.com/scholzhannah/CWComplexes/blob/7be4872a05b534011cc969eb5b80a4b7f0bf57e2/CWcomplexes/Finite.lean#L28-L43}{\faExternalLink}

\begin{lstlisting}
class FiniteDimensional.{u} {X : Type u} [TopologicalSpace X] (C : Set X) [CWComplex C] : Prop where
  eventually_isEmpty_cell : ∀ᶠ n in Filter.atTop, IsEmpty (cell C n)

class FiniteType.{u} {X : Type u} [TopologicalSpace X] (C : Set X) [CWComplex C] : Prop where
  finite_cell (n : ℕ) : Finite (cell C n)

class Finite.{u} {X : Type u} [TopologicalSpace X] (C : Set X) [CWComplex C] : Prop where
  eventually_isEmpty_cell : ∀ᶠ n in Filter.atTop, IsEmpty (cell C n)
  finite_cell (n : ℕ) : Finite (cell C n)
\end{lstlisting}

Property \lstinline{eventually_isEmpty_cell} is stated in terms of a \emph{filter}
\href{https://github.com/leanprover-community/mathlib4/blob/ed125a4216d18273cb1b96d4c846d32b85d74faf/Mathlib/Order/Filter/Basic.lean#L85-L96}{\faExternalLink}
which is a concept that appears frequently in mathlib.
They are often used to describe convergence in a topological way.
As they will not be important to this thesis we will not go into details but information on filters can be found in \cite{Bourbaki1966}.
The property \lstinline{eventually_isEmpty_cell} is equivalent to \lstinline{∃ a, ∀ (b : ℕ), a ≤ b → IsEmpty (cell C b)}.

To finish of this section here are the statements of some of the main results of Section \ref{sec:mathsdef}.
They correspond to the results \ref{lem:skeletonunionopenCell}
\href{https://github.com/scholzhannah/CWComplexes/blob/7be4872a05b534011cc969eb5b80a4b7f0bf57e2/CWcomplexes/Definition.lean#L320-L322}{\faExternalLink},
 \ref{lem:Frontiersubsetopen}
\href{https://github.com/scholzhannah/CWComplexes/blob/7be4872a05b534011cc969eb5b80a4b7f0bf57e2/CWcomplexes/Definition.lean#L420-L452}{\faExternalLink},
 \ref{lem:discretelevel0}
\href{https://github.com/scholzhannah/CWComplexes/blob/7be4872a05b534011cc969eb5b80a4b7f0bf57e2/CWcomplexes/Lemmas.lean#L108-L109}{\faExternalLink}, 
\ref{lem:compactintersectsonlyfinite} 
\href{https://github.com/scholzhannah/CWComplexes/blob/7be4872a05b534011cc969eb5b80a4b7f0bf57e2/CWcomplexes/Lemmas.lean#L111-L214}{\faExternalLink}
and \ref{lem:finiteiffcompact}
\href{https://github.com/scholzhannah/CWComplexes/blob/7be4872a05b534011cc969eb5b80a4b7f0bf57e2/CWcomplexes/Lemmas.lean#L298-L299}{\faExternalLink}.

\begin{lstlisting}
lemma iUnion_openCell_eq_level (n : ℕ∞) :
  ⋃ (m : ℕ) (_ : m < n + 1) (j : cell C m), openCell m j = level C n := ⋯

lemma cellFrontier_subset_finite_openCell (n : ℕ) (i : cell C n) :
    ∃ I : Π m, Finset (cell C m), cellFrontier n i ⊆
    (⋃ (m < n) (j ∈ I m), openCell m j) := ⋯

lemma isDiscrete_level_zero {A : Set X} : IsClosed (A ∩ level C 0) := ⋯

lemma compact_inter_finite (A : Set X) (compact : IsCompact A) :
  _root_.Finite
  (Σ (m : ℕ), {j : cell C m // ¬ Disjoint A (openCell m j)}) := ⋯

lemma compact_iff_finite : IsCompact C ↔ Finite C := ⋯
\end{lstlisting}

Where \lstinline|{a : α // P a}| is the subtype
\href{https://github.com/leanprover/lean4/blob/0edf1bac392f7e2fe0266b28b51c498306363a84/src/Init/Prelude.lean#L575-L590}{\faExternalLink}
of \lstinline{α} of all \lstinline{a : α} such that \lstinline{P a} for \lstinline{P : α → Prop}.
