\documentclass{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{amssymb}
\usepackage{upgreek}
\usepackage{mathtools}
\usepackage{amsthm}


\input{thesis_macros}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}

\begin{document}

\begin{lemma}
    Let $C$ be a CW-complex in a Hausdorff space $X$ as in the definition in the formalisation.
    Then $C$ is a CW-complex as in the paper definition.
  \end{lemma}
  \begin{proof}
    Properties (i), (ii), (iii) and (v) of the definition are immediate.
    Thus let us look at property (iv).
    We assume that
    \[A \subseteq C \text{ is closed in } X \iff \closedCell{n}{i} \cap A \text{ is closed in } X \text{ for all } n \in \bN \text{ and } i \in I_n\]
    and need to show that
    \[A \subseteq C \text{ is closed in } C \iff \closedCell{n}{i} \cap A \text{ is closed in } C \text{ for all } n \in \bN \text{ and } i \in I_n.\]
    It is easy to see that the forward direction is true.
    For the backwards direction take $A \subseteq C$ such that $A \cap \closedCell{n}{i}$ is closed in $C$ for all $n \in \bN$ and $i \in I_n$.
    That means that for every $n \in \bN$ and $i \in I_n$ there is a closed set $B_i^n \subseteq X$ such that $B^n_i \cap C = A \cap \closedCell{n}{i}$.
    But since $C$ is closed that means that $A \cap \closedCell{n}{i}$ was already closed for every $n \in \bN$ and $i \in I_n$.
    Thus we are done by assumption.
  \end{proof}


  \begin{lemma}\label{lem:weaktopologyproduct}
    If $X \times Y$ is a k-space then it has weak topology,
    i.e. $A \subseteq X \times Y$ is closed iff $\closedCell{n}{i} \times \closedCellf{m}{j} \cap A$ is closed for all $n, m \in \bN$, $i \in I_n$ and $j \in J_m$. 
\end{lemma}
\begin{proof}
    The forward direction is easy.

    Moving on to the backward direction we know that $A$ is closed if for every compact set $C \subseteq (X \times Y)_c$, $A \cap C$ is closed in $C$.
    Take such a compact set $C$.
    The projections $\pr{1}{C}$ and $\pr{2}{C}$ are compact as images of a compact set. 
    There are finite sets $E \subseteq \{e_i^n \mid n \in \bN, i \in I_n \}$ and $F \subseteq \{f_j^m \mid m \in \bN, j \in J_m \}$ s.t $\pr{1}{C} \subseteq \bigcup_{e \in E} e$ and $\pr{2}{C} \subseteq \bigcup_{f \in F} f$.
    Thus 
    \[C \subseteq \pr{1}{C} \times \pr{2}{C} \subseteq \bigcup_{e \in E} e \times \bigcup_{f \in F} f = \bigcup_{e \in E} \bigcup_{f \in F} e \times f.\] 
    So $C$ is included in a finite union of cells of $(X \times Y)_c$. 
    Therefore 
    \[A \cap C = A \cap \left (\bigcup_{e \in E} \bigcup_{f \in F} e \times f \right )\cap C = \left (\bigcup_{e \in E} \bigcup_{f \in F} A \cap (e \times f)\right ) \cap C\] 
    is closed since by assumption $A \cap (e \times f)$ is closed for every $e$ and $f$ and the union is finite. Thus $A \cap C$ is in particular closed in $C$.
\end{proof}

\end{document}