\documentclass{beamer}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{amssymb}
\usepackage{upgreek}
\usepackage{mathtools}

\usepackage{color}
%colors for lean syntax
\definecolor{keywordcolor}{rgb}{0.7, 0.1, 0.1}   % red
\definecolor{tacticcolor}{rgb}{0.0, 0.1, 0.6}    % blue
\definecolor{commentcolor}{rgb}{0.4, 0.4, 0.4}   % grey
\definecolor{symbolcolor}{rgb}{0.0, 0.1, 0.6}    % blue
\definecolor{sortcolor}{rgb}{0.1, 0.5, 0.1}      % green
\definecolor{attributecolor}{rgb}{0.7, 0.1, 0.1} % red

\def\lstlanguagefiles{lstlean.tex}
% set default language
\lstset{language=lean}

\usetheme{Madrid}
\usecolortheme{default}

\input{thesis_macros}

%Information to be included in the title page:
\title{Formalisation of CW-complexes}
\author{Hannah Scholz}
\institute[MI]{Mathematical Institute of the University of Bonn}
\date{20.09.2024}



\begin{document}

\frame{\titlepage}

\begin{frame}
\frametitle{Introduction: Lean, mathlib and CW-complexes}
\begin{itemize}
  \item Lean: Functional programming language and popular theorem prover
  \item Mathlib: Leans mathematical library with over 300 contributors and 28 maintainers
  \item CW-complexes: important concept in topology invented by Whitehead to state and prove the Whitehead theorem
  \item CW-complexes are not yet in mathlib 
  \item[$\blacktriangleright$] Aim: Provide a basic theory of CW-complexes and potentially contribute it to mathlib
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Definition of CW-complexes}
  Let $X$ be a Hausdorff space.
    A \emph{CW-complex} on $X$ consists of a family of indexing sets $(I_n)_{n \in \bN}$ and a family of maps $(Q_i^n\colon D_i^n\rightarrow X)_{n \ge 0, i \in I_n}$ s.t.
    \setbeamertemplate{enumerate items}[default]
    \begin{enumerate}[(i)]
        \item $\restrict{Q_i^n}{\interior{D_i^n}}\colon \interior{D_i^n} \rightarrow Q_i^n(\interior{D_i^n})$ is a homeomorphism. We call $\openCell{n}{i} \coloneq Q_i^n(\interior{D_i^n})$ an \emph{(open) $n$-cell} (or a cell of dimension $n$)
        and $\closedCell{n}{i} \coloneq Q_i^n(D_i^n)$ a \emph{closed $n$-cell}.
        \item For all $n, m \in \bN$, $i \in I_n$ and $j \in I_m$ where $(n, i) \ne (m, j)$ the cells $\openCell{n}{i}$ and $\openCell{m}{j}$ are disjoint.
        \item For each $n \in \bN$, $i \in I_n$, $Q_i^n(\boundary D_i^n)$ is contained in the union of a finite number of closed cells of dimension less than $n$.
        \item $A \subseteq X$ is closed iff $Q_i^n(D_i^n) \cap A$ is closed for all $n \in \bN$ and $i \in I_n$.
        \item $\bigcup_{n \ge 0}\bigcup_{i \in I_n} Q_i^n(D_i^n) = X$.
    \end{enumerate}
    We call $Q_i^n$ a \emph{characteristic map} and $\cellFrontier{n}{i} \coloneq Q_i^n(\boundary D_i^n)$ the \emph{frontier of the $n$-cell} for any $i$ and $n$.
    Additionally we define $X_n \coloneq \bigcup_{m < n + 1} \bigcup_{i \in I_m} \closedCell{m}{i}$ and call it the \emph{$n$-skeleton} of $X$ for $-1 \le n \le \infty$.
\end{frame}

\begin{frame}[fragile]
\frametitle{Lean: Definition of CW-complexes}
\begin{lstlisting}[basicstyle=\ttfamily\footnotesize]
class CWComplex.{u} {X : Type u} [TopologicalSpace X] (C : Set X) where
  cell (n : ℕ) : Type u
  map (n : ℕ) (i : cell n) : PartialEquiv (Fin n → ℝ) X
  source_eq (n : ℕ) (i : cell n) : (map n i).source = closedBall 0 1
  cont (n : ℕ) (i : cell n) : ContinuousOn (map n i) (closedBall 0 1)
  cont_symm (n : ℕ) (i : cell n) : ContinuousOn (map n i).symm (map n i).target
  pairwiseDisjoint' :
    (univ : Set (Σ n, cell n)).PairwiseDisjoint (fun ni ↦ map ni.1 ni.2 '' ball 0 1)
  mapsto (n : ℕ) (i : cell n) : ∃ I : Π m, Finset (cell m),
    MapsTo (map n i) (sphere 0 1) (⋃ (m < n) (j ∈ I m), map m j '' closedBall 0 1)
  closed' (A : Set X) (asubc : A ⊆ C) : IsClosed A ↔ ∀ n j, IsClosed (A ∩ map n j '' closedBall 0 1)
  union' : ⋃ (n : ℕ) (j : cell n), map n j '' closedBall 0 1 = C
\end{lstlisting}
\end{frame}

\begin{frame}
\frametitle{Equivalence of the definitions}
\begin{block}{Lemma}
  Let $C$ be a CW-complex in a Hausdorff space $X$ as in the definition in the formalisation.
  Then $C$ is a CW-complex as in the paper definition.
\end{block}
\begin{block}{Lemma}
  Let $X$ be a Hausdorff space and and $C$ a CW-complex in $X$ as in the formalised definition.
  Then $C$ is closed.
\end{block}
\end{frame}

\begin{frame}[fragile]{Closedness of CW-complexes}
  \begin{exampleblock}{Proof}
    \small
    \begin{tabular}{ll}
      1& By the weak topology it is enough to show that the intersection with every \\
      & closed cell is closed.\\
      2& Take any closed cell of $C$.\\
      3& Since the closed cell is a subset of $C$, the intersection is just the closed cell.\\
      4& Every closed cell is closed.\\
    \end{tabular}
  \end{exampleblock}
  \begin{exampleblock}{Proof (Lean)}
\begin{lstlisting}[basicstyle=\ttfamily\small, numbers=left, xleftmargin=21pt]
rw [closed _ (by rfl)]
intros
rw [inter_eq_right.2 (closedCell_subset_complex _ _)]
exact isClosed_closedCell
\end{lstlisting}
  \end{exampleblock}

\end{frame}

\begin{frame}{The topology of the product}
\begin{block}{Definition k-space}
  Let $X$ be a topological space. 
  We call $X$ a \emph{k-space} if 
  \begin{align*}
    A \subseteq X \text{ is closed} \iff &\text{for all compact sets } C \subseteq X \text{ the intersection } A \cap C \\
    &\text{is closed in } C.
  \end{align*}
\end{block}
\begin{block}{Lemma}
  Let $X$ be a CW-complex and $C \subseteq X$ a compact set. 
  Then $C$ is disjoint with all but finitely many cells of $X$.
\end{block}
\begin{block}{Lemma}
  If $X \times Y$ is a k-space then it has weak topology with respect to the characteristic maps $(Q_i^n \times P_j^m \colon D_i^n \times D_j^m \to (X \times Y)_c)_{n,m \in \bN,i \in I_n,j \in J_m}$, i.e. $A \subseteq X \times Y$ is closed iff $A \cap (\closedCell{n}{i} \times \closedCellf{m}{j})$ is closed for every $n, m \in \bN$, $i \in I_n$ and $j \in J_m$.
\end{block}
\end{frame}

\end{document}