\section{Basic constructions}

\todo[plan]{
I am not sure if this section should even exist.    
But I could briefly talk about: 
  \begin{enumerate}
    \item attaching cells
    \item disjoint unions?
    \item transporting along partial homeomorphisms?
  \end{enumerate}}

We have formalized a handful of miscellaneous constructions that we present in this section.

Firstly, we provide a way to get CW complex instances for non-definitional equalities. 
In Lean, this is 

\begin{lstlisting}[frame=single]
def RelCWComplex.ofEq {X : Type*} [TopologicalSpace X] (C D : Set X)
    {E F : Set X} [RelCWComplex C D] (hCE : C = E) (hDF : D = F) : 
    RelCWComplex E F :=
  /- Proof omitted-/
\end{lstlisting}

We also want to be able to transport CW-structures along homeomorphisms. 
Because of our local approach, the more general statement features ``local homeomorphisms'': 

\begin{lstlisting}[frame=single]
def RelCWComplex.ofPartialEquiv.{u} {X Y : Type u} [TopologicalSpace X] 
    [T2Space X] [TopologicalSpace Y] (C : Set X) {D : Set X} (E : Set Y) 
    {F : Set Y} [RelCWComplex C D] (hE : IsClosed E) (f : PartialEquiv X Y) (hfC1 : f.source = C) (hfE1 : f.target = E) (hDF : f '' D = F) 
    (hfC2 : ContinuousOn f C) (hfE2 : ContinuousOn f.symm E) :
    RelCWComplex E F := 
  /- Proof omitted-/
\end{lstlisting}

We can then easily derive a version for the general ``total'' homeomorphisms.

Next, we show that the disjoint union of two CW complexes is again a CW complex. 
Here, we again simplify the problem by considering both CW complexes as subsets of the same topological space. 
The sufficient conditions for this to work are that the complexes are disjoint and their bases are separated by neighbourhoods. 
This translates to: 

\begin{lstlisting}[frame=single]
def RelCWComplex.disjointUnion [RelCWComplex C D] {E F : Set X} 
    [RelCWComplex E F] (hCE : Disjoint C E) (hDF : SeparatedNhds D F) : RelCWComplex (C ∪ E) (D ∪ F) :=
  /- Proof omitted-/
\end{lstlisting}

\todo[plan]{Here, I would like to say something about this not being an instance and us therefore providing one but I have not managed to define this instance :(}

We provide definitions that allow you to attach one cell at a time or attach a collection of cells of the same dimension at once.
These are in convenient when constructing examples of CW complexes. 
Lastly, we make it possible to restrict and enlarge the ambient type around a CW complex.
